---
title: "StarbucksHistoryInvestigation"
author: "Ruinan(Victor) Zhang"
date: "11/9/2016"
output: word_document
---

# Background
This task is to investigate growth of chain companies such as Starbucks. The approach is to construct a dynamic graphic to illustrate Starbuck stores' expension. The date is provided by Winter who setted his mission drink coffeefrom every company-owned (including partner companies) Starbucks in the world. The openning dates of the starbucks are provided on [his website](http://www.starbuckseverywhere.net/StoreOpeningDates.htm): 

# Approach 
To download the data from Winter's website, I used the web scrapping tools from rvest package to download and clean the informations about every starbucks stores he visited by 11/03/2016. 
```{r,eval=FALSE}
# read hdml
url = "http://www.starbuckseverywhere.net/StoreOpeningDates.htm"
html <- html(url)

# retrieve data from Winter's website and save it
df <- html %>% html_nodes("table") %>% html_table()
save(df,file = "starbucks_data.Rda")
```

Based on the address of each store, I used geocode() function to request the longtitude and latitude of every store. 
```{r,eval=FALSE}
geo_locations <- data.frame()  # df for storing all the geo_locations of starbucks 
for(i in 1:nrow(df)){
  row <- df[i,]
  location_str <- paste(row$Market,row$City,row$Name)  # location str
  geo<-geocode(location_str,source = "dsk")            # call geo_code() to get locations
  geo_locations<- rbind(geo_locations,geo)
}
geo_locations
```

After achieving the locations in longitude and latitude, I sorted all the stores by there openning date and construct a dynamic graphic using the "animation" package and stored the output as a gif file. 
```{r,eval=FALSE}
# sort df by year, month, and data
date_info <- as.data.frame(matrix(date_info,ncol=3,byrow = T))
names(date_info) <- c("month",'day','year')
df <- cbind(df,date_info)
df <- (df[order(df$year,df$month,df$day),])

# drop cases with NA for locations
geo_locations = geo_locations[complete.cases(geo_locations), ]

# get the base satellite map of US
mapgilbert <- get_map(location = c(lon = -98.5795, lat = 39.8282), zoom = 4,
                      maptype = "satellite", scale = 2)
g<-ggmap(mapgilbert)+ggtitle("Starbucks in US")+guides(fill=FALSE)+xlab("longitude")+ylab("latitude")
g

# plot function for the animation
plotfoo<- function(){
  i=1
  while(i<nrow(geo_locations)){
    # the increment is dynamic
    increment = ceiling(i/2)
    if(increment>50){
      increment=50
    }
    bound = i+increment
    if(bound>nrow(geo_locations)){
      bound = nrow(geo_locations)
    }
    each_location <- geo_locations[i:bound,]   # for each iteraton, plot i to i+increment rows
    print(i)
    i=i+increment
    g <- g+geom_point(data=each_location,aes(x = lon, y = lat, fill = "red"), size = 2, shape = 21,alpha=0.8)
    g <- g+ggtitle(paste("Starbucks in US by",geo_locations[i,]$year))
    print(g)
  }
}

saveGIF(plotfoo(),interval = 0.1,movie.name = "starbucks.gif")
```


# Result
Based on the animation, it is clear that majority of the starbucks are opened in large cities. Starbucks is an organization initiated from Seattle in Washington state and quickly expend to both east and west coasts. The growth of starbucks really bursted from 2006 to 2008. However, for state like Wisconsin and Kansas, it seems like Starbucks have very fiew stores in those areas. There are definitely many stores in most area of east cost and midwest and major cities on west coast. 

The dynamic graph is attached as a seperate gif file. 

# Full Code
```{r,eval=FALSE}
library(ggmap)
library(ggplot2)
library(rvest)
library(dplyr)
library(maps)
library(animation)
library(stringr)

url = "http://www.starbuckseverywhere.net/StoreOpeningDates.htm"

html <- html(url)

# retrieve data from Winter's website
df <- html %>% html_nodes("table") %>% html_table()
save(df,file = "starbucks_data.Rda")
 load("starbucks_data.Rda")

 df <- df[[1]]
 colnames(df) <- df[1,]
 df <- df[2:nrow(df),]
 # clean backslahes in teh Market column
 df$Market <-str_replace(df$Market,"\\\\"," ")


row = df[1,]
# clean out the date infos
date_info<- str_split(df$Opened,"/")
date_info <- unlist(date_info)
date_info <- as.numeric(date_info)

# sort df by year, month, and data
date_info <- as.data.frame(matrix(date_info,ncol=3,byrow = T))
names(date_info) <- c("month",'day','year')
df <- cbind(df,date_info)
df <- (df[order(df$year,df$month,df$day),])


geo_locations <- data.frame()  # df for storing all the geo_locations of starbucks 
for(i in 1:nrow(df)){
  row <- df[i,]
  location_str <- paste(row$Market,row$City,row$Name)
  geo<-geocode(location_str,source = "dsk")
  geo_locations<- rbind(geo_locations,geo)
}
geo_locations <- cbind(df,geo_locations)

# save the geo_locations
save(geo_locations,file = "geo_locations.Rda")
load("geo_locations.Rda")

# drop cases with NA for locations
geo_locations = geo_locations[complete.cases(geo_locations), ]

# get the base satellite map of US
mapgilbert <- get_map(location = c(lon = -98.5795, lat = 39.8282), zoom = 4,
                      maptype = "satellite", scale = 2)
g<-ggmap(mapgilbert)+ggtitle("Starbucks in US")+guides(fill=FALSE)+xlab("longitude")+ylab("latitude")
g

# plot function for the animation
plotfoo<- function(){
  i=1
  while(i<nrow(geo_locations)){
    # the increment is dynamic
    increment = ceiling(i/2)
    if(increment>50){
      increment=50
    }
    bound = i+increment
    if(bound>nrow(geo_locations)){
      bound = nrow(geo_locations)
    }
    each_location <- geo_locations[i:bound,]   # for each iteraton, plot i to i+increment rows
    print(i)
    i=i+increment
    g <- g+geom_point(data=each_location,aes(x = lon, y = lat, fill = "red"), size = 2, shape = 21,alpha=0.8)
    g <- g+ggtitle(paste("Starbucks in US by",geo_locations[i,]$year))
    print(g)
  }
}

saveGIF(plotfoo(),interval = 0.1,movie.name = "starbucks.gif")




```


